---
title: "\\vspace{-1.5cm}{WGS Quality Control Report}"
output:
  pdf_document: 
    latex_engine: xelatex
    keep_tex: true
mainfont: Helvetica
papersize: a4
header-includes:
- \usepackage{titling}
- \pretitle{\begin{flushleft}}
- \posttitle{\end{flushleft}} 
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{makecell}
- \usepackage{xcolor}
- \usepackage{soul}
- \usepackage{caption}
- \usepackage[singlelinecheck=false]{caption}
- \usepackage[font={footnotesize,it}]{caption}

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load library, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(readxl)
library(dplyr)
library(kableExtra)
```

```{r call data, echo=FALSE, message=FALSE, warning=FALSE}
setwd("D:/ALLYSA FILE/2023/DMU Files/DMU Projects/WGS_QC")
wgs_df <- read.csv("data_files/qualifyr_report_2023-08-09.csv")
```


```{r rename columns, echo=FALSE, message=FALSE, warning=FALSE}
wgs_df <- wgs_df %>% 
  rename(
    sample_id = sample_name,
    wgs_id = bactinspector.species.metric_value,
    check_result = bactinspector.result.check_result,
    contamination_value = confindr.percentage_contamination.metric_value,
    contamination_result =confindr.percentage_contamination.check_result,
    no_of_contigs_value = quast...contigs.....1000.bp..metric_value,
    no_of_contigs_result = quast...contigs.....1000.bp..check_result,
    gc_percent_value = quast.GC.....metric_value,
    gc_percent_result = quast.GC.....check_result,
    n50_value = quast.N50.metric_value,
    n50_result = quast.N50.check_result,
    total_length_value = quast.Total.length.....1000.bp..metric_value,
    total_length_result = quast.Total.length.....1000.bp..check_result
    )

#write.csv(wgs_df, file="wgs_qc_data.csv")
```


```{r code organism names, echo=FALSE, message=FALSE, warning=FALSE}
org_df <- read_xlsx("data_files/ORGLIST.xlsx")
org_df <- subset(org_df, select = c(ORG,ORGANISM))

result_df <- org_df %>% 
  right_join(wgs_df, by= c('ORGANISM' = 'wgs_id'))

result_df <- subset(result_df, select = c('ORG', 'sample_id'))
wgs_df <- merge(x=wgs_df,y=result_df, by="sample_id", all.x=TRUE)

wgs_df$ORG[grepl("Salmonella",wgs_df$wgs_id)] = "sal"
wgs_df$ORG[grepl("Shigella",wgs_df$wgs_id)] = "shi"
wgs_df$ORG[grepl("Vibrio",wgs_df$wgs_id)] = "vic"

wgs_df$warning_count <- ifelse(wgs_df$check_result == 'WARNING',1,0)

#write.csv(wgs_df, file="wgs_qc_data.csv")
```


\fontsize{7}{8}
\selectfont
\captionsetup[table]{labelformat=empty}
\renewcommand{\arraystretch}{1.2}

```{r qc listing, echo=FALSE, message=FALSE, warning=FALSE}
arsrl_result_df <- read.csv("data_files/arsrl_result_2023-08-09.csv")

retain_column <- c('sample_id','wgs_id')
wgs_result_df <- subset(wgs_df, select = retain_column)

qc_listing_df <- merge(x=arsrl_result_df,y=wgs_result_df, by="sample_id", all.x=TRUE)

qc_listing_nc <- qc_listing_df %>%
  rowwise() %>%
  mutate(concordant = 
           grepl(arsrl_result, wgs_id)
         )

warning_color <- which(wgs_df$result == 'WARNING')
nc_warning <- ifelse(qc_listing_nc$concordant == TRUE, "white","#ffcbd1")

colnames(qc_listing_df) <- c('Sample ID','ARSRL','WGS') 

qc_listing_df %>% 
  mutate(WGS = cell_spec(WGS, background = nc_warning)) %>%
  kable(booktabs = T, align = "c", escape = FALSE)%>%
  column_spec(1, width = "2cm") %>%
  column_spec(2, width = "5cm") %>%
  column_spec(3, width = "7cm") %>%
  row_spec(warning_color, background = "#F9E4BC")%>%
  row_spec(0, background = "#D4D4D4",bold = TRUE,align = "c")

```

\tiny Legend:
```{r qc listing legend, echo=FALSE, message=FALSE, warning=FALSE,results='asis'}
lgd1 <- c("PASS")
lgd2 <- c("WARNING")
lgd3 <- c("EXCEEDS THRESHOLD METRIC/S")
lgd4 <- c("NON-CONCORDANT")
  
legend_tbl <- data.frame(lgd1,lgd2,lgd3,lgd4)

legend_tbl[1,1] = cell_spec(legend_tbl[1,1],font_size = 4)
legend_tbl[1,2] = cell_spec(legend_tbl[1,2],font_size = 4)
legend_tbl[1,3] = cell_spec(legend_tbl[1,3],font_size = 4)
legend_tbl[1,4] = cell_spec(legend_tbl[1,4],font_size = 4)


legend_tbl %>%
  kable(booktabs = T, align = "c", escape = FALSE, col.names = NULL) %>%
  column_spec(1, width = "1cm",border_left = TRUE,background = "white")%>%
  column_spec(2, width = "1cm",border_left = TRUE, background = "#F9E4BC")%>%
  column_spec(3, width = "3cm",border_left = TRUE,color = "red")%>%
  column_spec(4, width = "3cm",background = "#ffcbd1",border_left = TRUE, border_right =  TRUE)

#write.csv(wgs_df, file="wgs_qc_data.csv")
```

$\\$

\fontsize{7}{8}
\selectfont
\captionsetup[table]{labelformat=empty}
\renewcommand{\arraystretch}{1.2}
```{r summary table, echo=FALSE, message=FALSE, warning=FALSE}
retain_column <- c('sample_id','contamination_value','no_of_contigs_value','gc_percent_value','n50_value','total_length_value')
wgs_summary_df <- subset(wgs_df, select = retain_column)

wgs_summary_df$sample_id <- ifelse(wgs_df$warning_count == 1,paste("*",wgs_df$sample_id),wgs_df$sample_id)
warning_color <- which(wgs_df$result == 'WARNING')
contigs_warning <- ifelse(wgs_df$no_of_contigs_result == 'WARNING', "red","black")
n50_warning <- ifelse(wgs_df$n50_result == 'WARNING', "red","black")

colnames(wgs_summary_df) <- c('SAMPLE ID','CONTAMINATION','CONTIGS', 'GC PERCENT','N50','TOTAL LENGTH') 

wgs_summary_df %>% 
  mutate(N50 = cell_spec(N50, color = n50_warning)) %>%
  mutate(CONTIGS = cell_spec(CONTIGS, color = contigs_warning)) %>%
  kable(booktabs = T, align = "c", escape = FALSE) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(2, width = "2cm") %>%
  column_spec(3, width = "3cm") %>%
  column_spec(4, width = "2cm") %>%
  column_spec(5, width = "2cm") %>%
  column_spec(6, width = "2cm") %>%
  row_spec(warning_color, background = "#F9E4BC") %>%
  row_spec(0, background = "#D4D4D4",bold = TRUE,align = "c")

#write.csv(wgs_df, file="wgs_qc_data.csv")
```


\tiny Legend:
```{r summary table legend, echo=FALSE, message=FALSE, warning=FALSE,results='asis'}
lgd1 <- c("PASS")
lgd2 <- c("WARNING")
lgd3 <- c("EXCEEDS THRESHOLD METRIC/S")
lgd4 <- c("*Isolates were tagged with warning due to uncertain results of species identification using bactinspector")
  
legend_tbl <- data.frame(lgd1,lgd2,lgd3,lgd4)

legend_tbl[1,1] = cell_spec(legend_tbl[1,1],font_size = 4)
legend_tbl[1,2] = cell_spec(legend_tbl[1,2],font_size = 4)
legend_tbl[1,3] = cell_spec(legend_tbl[1,3],font_size = 4)
legend_tbl[1,4] = cell_spec(legend_tbl[1,4],font_size = 4)


legend_tbl %>%
  kable(booktabs = T, align = "c", escape = FALSE, col.names = NULL) %>%
  column_spec(1, width = "1cm",border_left = TRUE,background = "white")%>%
  column_spec(2, width = "1cm",border_left = TRUE, background = "#F9E4BC")%>%
  column_spec(3, width = "2.5cm",border_left = TRUE,color = "red")%>%
  column_spec(4, width = "6.5cm",border_left = TRUE, border_right =  TRUE)

#write.csv(wgs_df, file="wgs_qc_data.csv")
```

\fontsize{7}{8}
\selectfont
\captionsetup[table]{labelformat=empty}
\renewcommand{\arraystretch}{1.2}


```{r qc listing summary narrative, echo=FALSE, message=FALSE, warning=FALSE}
qc_summary_df <- qc_listing_df %>% group_by(WGS) %>% summarise(number = sum(!is.na(WGS)))
qc_summary_df <- qc_summary_df[order(-qc_summary_df$number),]

qc_tl <- nrow(qc_listing_df)
qc_distinct <- nrow(qc_summary_df)

qc_pass <- nrow(wgs_df[wgs_df$result == 'PASS', ])
qc_warning <- nrow(wgs_df[wgs_df$result == 'WARNING', ])

qc_pass_pc <- if (qc_pass != 0) (qc_pass/qc_tl) * 100 else 0
qc_pass_pc <- format(round(qc_pass_pc, 2), nsmall = 2)

qc_warning_pc <- if (qc_warning != 0) (qc_warning/qc_tl) * 100 else 0
qc_warning_pc <- format(round(qc_warning_pc, 2), nsmall = 2)

qc_concordance <- nrow(qc_listing_nc[qc_listing_nc$concordant == TRUE, ])
qc_cn_pc <- if (qc_concordance != 0) (qc_concordance/qc_tl) * 100 else 0
qc_cn_pc <- format(round(qc_cn_pc, 2), nsmall = 2)

```

$\\$
$\\$

```{r qc listing summary table, echo=FALSE, message=FALSE, warning=FALSE}
colnames(qc_summary_df) <- c('WGS_ID','Number')

qc_summary_df %>% 
  kable(booktabs = T, align = "lc")%>%
  column_spec(1, width = "8cm") %>%
  row_spec(0, background = "#D4D4D4",bold = TRUE)

```
* $\color{red}`r qc_distinct`$ distinct species were identified among $\color{red}`r qc_tl`$
isolates.

* $\color{red}`r qc_pass_pc`$ % of the isolates passed the QC, while $\color{red}`r qc_warning_pc`$ % were tagged with warning.


* Concordance between ARSRL and WGS species report was $\color{red}`r qc_cn_pc`$ %.


$\\$

### GRAPHS

\fontsize{7}{8}
\selectfont
\captionsetup[table]{labelformat=empty}
\renewcommand{\arraystretch}{1.2}


```{r organism summary, echo=FALSE, message=FALSE, warning=FALSE}
# Calculate the counts for each "wgs_id"
wgs_id_count <- table(wgs_df$wgs_id)

# Create a data frame with "wgs_id" and their counts
wgs_count_df <- data.frame(wgs_id = names(wgs_id_count), Number = as.vector(wgs_id_count))

# Sort the data frame by the "Number" column in descending order
wgs_count_df <- wgs_count_df[order(-wgs_count_df$Number),]

# Now, set the factor levels in reverse order for correct sorting
wgs_count_df$wgs_id <- factor(wgs_count_df$wgs_id, levels = rev(wgs_count_df$wgs_id))

# Create the bar plot using ggplot with labels
ggplot(data = wgs_count_df) +
  geom_bar(mapping = aes(x = wgs_id, y = Number), stat = "identity", width = 0.5, position = position_dodge(width = 1),fill="lightblue") +  
  coord_flip() +
  geom_text(aes(x = wgs_id, y = Number, label = Number), hjust = -0.1, size = 4, color = "black") +
  labs(y = "Number")+
  theme(text = element_text(size=8))


```


### Result Classification
```{r organism results, echo=FALSE, message=FALSE, warning=FALSE}
# Define the desired order for the 'result' levels
desired_order <- c("PASS", "WARNING", "FAILURE")
wgs_df$result <- factor(wgs_df$result, levels = desired_order)

result_colors <- c("FAILURE" = "red", "PASS" = "green", "WARNING" = "purple")

ggplot(data = wgs_df, aes(x = ORG, fill = result)) +
  geom_bar() +
  facet_wrap(~ result) +
  coord_flip() +
  scale_fill_manual(values = result_colors) + # Apply the defined colors
  guides(fill = FALSE) #+  # Hide the legend for the 'result' variable
  #theme_minimal()
```


### Number of contigs
```{r ,echo=FALSE, message=FALSE, warning=FALSE}

ggplot(data = wgs_df) +
  geom_point(mapping = aes(x = no_of_contigs_value, y = ORG, color = ifelse(no_of_contigs_value < 301, "blue", "red"))) +
  scale_color_manual(values = c("blue", "red"), labels = c("Passed", "Above threshold")) +
  labs(color = "Contigs") +
  theme_minimal()
```



### List of samples above QC threshold metrics
```{r QC failed list,echo=FALSE, message=FALSE, warning=FALSE}
QC_failed <- wgs_df %>%
  filter(result == "FAILURE") %>%
  select(sample_id,  result, contamination_result, no_of_contigs_value, n50_value, total_length_value)

if (nrow(QC_failed) > 0) {
  kable(QC_failed)
} else {
  cat("No QC failures found.")
}
```
