---
#title: "\\vspace{-1.5cm}{WGS Quality Control Report}"
#subtitle: "\\fontsize{6pt}{2pt}{Batch name: 2024-01-25}"
title: \vspace{-1.5cm} \begin{LARGE} WGS Quality Control Report \end{LARGE}
subtitle: \begin{normalsize} Batch name:2024-01-25 \end{normalsize}
output:
  pdf_document: 
    latex_engine: xelatex
    keep_tex: true
mainfont: Helvetica
papersize: a4
header-includes:
- \usepackage{titling}
- \pretitle{\begin{flushleft}}
- \posttitle{\end{flushleft}} 
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{makecell}
- \usepackage{xcolor}
- \usepackage{soul}
- \usepackage{caption}
- \usepackage[singlelinecheck=false]{caption}
- \usepackage[font={footnotesize,it}]{caption}

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load library, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(readxl)
library(dplyr)
library(kableExtra)
library(RColorBrewer)
library(scales)
library(readr)
library(DBI)
library(svDialogs)

```

```{r call data, echo=FALSE, message=FALSE, warning=FALSE}
#get_file <- dlgInput("Enter filename", Sys.info()[" "])$res
#get_file <- paste("data_files/",get_file,".tsv", sep="")
#wgs_df <- read.delim(file= get_file)


setwd("D:/ALLYSA FILE/2023/DMU Files/DMU Projects/WGS_QC")
wgs_df <- read.delim(file="data_files/qualifyr_report_2024-01-25.tsv")

#wgs_df$sample_name <- gsub("_", "\\_\\", wgs_df$sample_name)
wgs_df$sample_name <- gsub("-", "_", wgs_df$sample_name, fixed=TRUE)

wgs_df[wgs_df == 'WARN'] <- 'WARNING'
wgs_df[wgs_df == 'FAIL'] <- 'FAILURE'
```


```{r rename columns, echo=FALSE, message=FALSE, warning=FALSE}
wgs_df <- wgs_df %>% 
  rename(
    sample_id = sample_name,
    wgs_id = bactinspector.species.metric_value,
    check_result = bactinspector.result.check_result,
    dups_result = fastqc.1.Sequence.Duplication.Levels.check_result,
    contamination_value = confindr.percentage_contamination.metric_value,
    contamination_result =confindr.percentage_contamination.check_result,
    no_of_contigs_value = quast...contigs.....0.bp..metric_value,
    no_of_contigs_result = quast...contigs.....0.bp..check_result,
    gc_percent_value = quast.GC.....metric_value,
    gc_percent_result = quast.GC.....check_result,
    n50_value = quast.N50.metric_value,
    n50_result = quast.N50.check_result,
    total_length_value = quast.Total.length.....1000.bp..metric_value,
    total_length_result = quast.Total.length.....1000.bp..check_result
    )

#write.csv(wgs_df, file="wgs_qc_data.csv")
```




```{r read ARSRL result, echo=FALSE, message=FALSE, warning=FALSE}
# taking input with showing the message
#get_file <- dlgInput("Enter a text filename", Sys.info()[" "])$res
id_list <- na.omit(wgs_df[['sample_id']])
id_list <- gsub("\\_", "_", id_list)

# Connect to db
con <- dbConnect(RPostgres::Postgres(),dbname = 'WGS_DB', 
                 host = '10.10.25.163', # i.e. 'ec2-54-83-201-96.compute-1.amazonaws.com'
                 port = 5432, # or any other port specified by your DBA
                 user = 'postgres',
                 password = 'secret123')


# Specify the path to your text file
#file_path <- paste('data_files/',get_file,'.txt', sep='') 

# Read the IDs from the text file
#id_list <- scan(file_path, what = "")

#list of query for txt file
#id_list <- scan("to_query.txt", character(), quote = "")

# Escape single quotes in the strings
id_list <- gsub("'", "''", id_list)

# Convert the list of IDs to a comma-separated string enclosed in single quotes
id_string <- paste0("'", paste(id_list, collapse = "','"), "'")


query <- paste("SELECT * from wgs_app_referreddb 
               WHERE wgs_app_referreddb.sample_name IN (", id_string, ")", sep="")

df <- dbSendQuery(con, query)

result <- dbFetch(df)
result <- result %>% mutate_all(as.character)
result <- result %>% mutate_all(~as.character(ifelse(. == "nan", "", .)))

result <- subset(result , select = c(sample_name,arsrl_org))
sample_name = c("UTP_BL_001","UTP_BL_002","UTP_BL_003","UTP_BL_004","UTP_BL_005","UTP_ST_003","UTP_ST_014")
arsrl_org = c("Escherichia coli","Escherichia coli","Escherichia coli","Escherichia coli","Escherichia coli","Escherichia coli","Escherichia coli")
#result <- rbind(result,sample_name,arsrl_org)

arsrl_result_df <- result %>% 
   add_row(sample_name = sample_name, arsrl_org=arsrl_org)


colnames(arsrl_result_df) <- c('sample_id','arsrl_org') 
```




```{r code organism names, echo=FALSE, message=FALSE, warning=FALSE}
org_df <- read_xlsx("data_files/ORGLIST.xlsx")
org_df <- subset(org_df, select = c(ORG,ORGANISM))

result_df <- org_df %>% 
  right_join(wgs_df, by= c('ORGANISM' = 'wgs_id'))

result_df <- subset(result_df, select = c('ORG', 'sample_id'))
wgs_df <- merge(x=wgs_df,y=result_df, by="sample_id", all.x=TRUE)

wgs_df$ORG[grepl("Salmonella",wgs_df$wgs_id)] = "sal"
wgs_df$ORG[grepl("Shigella",wgs_df$wgs_id)] = "shi"
wgs_df$ORG[grepl("Vibrio",wgs_df$wgs_id)] = "vic"

wgs_df$warning_count <- ifelse(wgs_df$check_result == 'WARNING',1,
                               ifelse(wgs_df$dups_result == 'WARNING',2,
                                      ifelse(wgs_df$check_result == 'WARNING' & wgs_df$dups_result == 'WARNING',3,0)))

wgs_warning_df <- filter(wgs_df, warning_count!=0 )
#write.csv(wgs_df, file="wgs_qc_data.csv")
```


\fontsize{7}{8}
\selectfont
\captionsetup[table]{labelformat=empty}
\renewcommand{\arraystretch}{1.2}

```{r qc listing, echo=FALSE, message=FALSE, warning=FALSE}
#arsrl_result_df <- read.csv("data_files/qualifyr_report_2024-01-29.csv")
#arsrl_result_df$sample_id <- gsub("_", "\\_\\", arsrl_result_df$sample_id)
arsrl_result_df$sample_id <- gsub("_", "\\_", arsrl_result_df$sample_id, fixed=TRUE)
wgs_df$sample_id <- gsub("_", "\\_", wgs_df$sample_id, fixed=TRUE)

retain_column <- c('sample_id','wgs_id')
wgs_result_df <- subset(wgs_df, select = retain_column)

qc_listing_df <- merge(x=arsrl_result_df,y=wgs_result_df, by="sample_id", all.x=TRUE)

# Define a function to get the part before the first white space
get_value_before_whitespace <- function(x) {
  parts <- strsplit(as.character(x), " ")[[1]]
  return(parts[1])
}

# Apply the function to the specified column
qc_listing_df$Genus <- sapply(qc_listing_df$arsrl_org, get_value_before_whitespace)

qc_listing_nc <- qc_listing_df %>%
  rowwise() %>%
  mutate(concordant = 
           grepl(Genus, wgs_id)
         )

warning_color <- which(wgs_df$result == 'WARNING')
failure_color <- which(wgs_df$result == 'FAILURE')
nc_warning <- ifelse(qc_listing_nc$concordant == TRUE,"white","yellow")

qc_listing_df = subset(qc_listing_df, select = -c(Genus) )

colnames(qc_listing_df) <- c('Sample ID','ARSRL','WGS')

qc_listing_df %>% 
  mutate(WGS = cell_spec(WGS, background = nc_warning)) %>%
  kable(booktabs = T, align = "c", escape = FALSE)%>%
  column_spec(1, width = "3cm") %>%
  column_spec(2, width = "5cm") %>%
  column_spec(3, width = "7cm") %>%
  row_spec(warning_color, background = "#FFA77F") %>%
  row_spec(failure_color, background = "#FD7979") %>%
  row_spec(0, background = "#D4D4D4",bold = TRUE,align = "c")

```


\tiny Legend:
```{r qc listing legend, echo=FALSE, message=FALSE, warning=FALSE,results='asis'}
lgd1 <- c("PASS")
lgd2 <- c("WARNING")
lgd3 <- c("FAILURE")
lgd4 <- c("EXCEEDS THRESHOLD METRIC/S")
lgd5 <- c("NON-CONCORDANT")

legend_tbl <- data.frame(lgd1,lgd2,lgd3,lgd4,lgd5)

legend_tbl[1,1] = cell_spec(legend_tbl[1,1],font_size = 4)
legend_tbl[1,2] = cell_spec(legend_tbl[1,2],font_size = 4)
legend_tbl[1,3] = cell_spec(legend_tbl[1,3],font_size = 4)
legend_tbl[1,4] = cell_spec(legend_tbl[1,4],font_size = 4)
legend_tbl[1,5] = cell_spec(legend_tbl[1,5],font_size = 4)


legend_tbl %>%
  kable(booktabs = T, align = "c", escape = FALSE, col.names = NULL) %>%
  column_spec(1, width = "1cm",border_left = TRUE,background = "white")%>%
  column_spec(2, width = "1cm",border_left = TRUE, background = "#FFA77F")%>%
  column_spec(3, width = "1cm",border_left = TRUE, background = "#FD7979")%>%
  column_spec(4, width = "3cm",border_left = TRUE,color = "blue")%>%
  column_spec(5, width = "3cm",background = "yellow",border_left = TRUE, border_right =  TRUE)



#write.csv(wgs_df, file="wgs_qc_data.csv")
```

$\\$

\fontsize{7}{8}
\selectfont
\captionsetup[table]{labelformat=empty}
\renewcommand{\arraystretch}{1.2}
```{r summary table, echo=FALSE, message=FALSE, warning=FALSE}
retain_column <- c('sample_id','contamination_value','no_of_contigs_value','gc_percent_value','n50_value','total_length_value')
wgs_summary_df <- subset(wgs_df, select = retain_column)

wgs_summary_df$sample_id <- ifelse(wgs_df$warning_count == 1,paste("*",wgs_df$sample_id),wgs_df$sample_id)
warning_color <- which(wgs_df$result == 'WARNING')
failure_color <- which(wgs_df$result == 'FAILURE')
contamination_warning <- ifelse(wgs_df$contamination_result == 'FAILURE', "blue",ifelse(wgs_df$contamination_result == 'WARNING', "blue","black"))
contigs_warning <- ifelse(wgs_df$no_of_contigs_result == 'FAILURE', "blue",ifelse(wgs_df$no_of_contigs_result == 'WARNING', "blue","black"))
n50_warning <- ifelse(wgs_df$n50_result == 'FAILURE', "blue",ifelse(wgs_df$n50_result == 'WARNING',"blue","black"))


colnames(wgs_summary_df) <- c('SAMPLE ID','CONTAMINATION','CONTIGS', 'GC PERCENT','N50','TOTAL LENGTH') 

wgs_summary_df %>% 
  mutate(CONTAMINATION = cell_spec(CONTAMINATION, color = contamination_warning)) %>%
  mutate(N50 = cell_spec(N50, color = n50_warning)) %>%
  mutate(CONTIGS = cell_spec(CONTIGS, color = contigs_warning)) %>%
  kable(booktabs = T, align = "c", escape = FALSE) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(2, width = "2cm") %>%
  column_spec(3, width = "3cm") %>%
  column_spec(4, width = "2cm") %>%
  column_spec(5, width = "2cm") %>%
  column_spec(6, width = "2cm") %>%
  row_spec(warning_color, background = "#FFA77F") %>%
  row_spec(failure_color, background = "#FD7979") %>%
  row_spec(0, background = "#D4D4D4",bold = TRUE,align = "c")

#write.csv(wgs_df, file="wgs_qc_data.csv")
```


\tiny Legend:
```{r summary table legend, echo=FALSE, message=FALSE, warning=FALSE,results='asis'}
warning_value <- unique(wgs_warning_df$warning_count)
warning_value <- ifelse(length(warning_value)==0,0,warning_value)


lgd1 <- c("PASS")
lgd2 <- c("WARNING")
lgd3 <- c("FAILURE")
lgd4 <- c("EXCEEDS THRESHOLD METRIC/S")

lgd5_value <- ifelse(warning_value == 1,"*Isolates were tagged with warning due to uncertain results of species identification using bactinspector.",ifelse(warning_value == 2,"*Isolates were tagged with warning due to uncertain results of sequence identification levels.",ifelse(warning_value == 3,"*Isolates were tagged with warning due to uncertain results of species identification using bactinspector and sequence identification levels.","*Isolates were tagged with warning due to uncertain results  of species identification using bactinspector or sequence identification levels.")))
  
  
lgd5 <- lgd5_value
#lgd4 <- c("*Isolates were tagged with warning due to uncertain results of species identification using bactinspector")
  
legend_tbl <- data.frame(lgd1,lgd2,lgd3,lgd4,lgd5)

legend_tbl[1,1] = cell_spec(legend_tbl[1,1],font_size = 4)
legend_tbl[1,2] = cell_spec(legend_tbl[1,2],font_size = 4)
legend_tbl[1,3] = cell_spec(legend_tbl[1,3],font_size = 4)
legend_tbl[1,4] = cell_spec(legend_tbl[1,4],font_size = 4)
legend_tbl[1,5] = cell_spec(legend_tbl[1,5],font_size = 4)


legend_tbl %>%
  kable(booktabs = T, align = "c", escape = FALSE, col.names = NULL) %>%
  column_spec(1, width = "1cm",border_left = TRUE,background = "white")%>%
  column_spec(2, width = "1cm",border_left = TRUE, background = "#FFA77F")%>%
  column_spec(3, width = "1cm",border_left = TRUE, background = "#FD7979")%>%
  column_spec(4, width = "2.5cm",border_left = TRUE,color = "blue")%>%
  column_spec(5, width = "8cm",border_left = TRUE, border_right =  TRUE)

#write.csv(wgs_df, file="wgs_qc_data.csv")
```

\fontsize{7}{8}
\selectfont
\captionsetup[table]{labelformat=empty}
\renewcommand{\arraystretch}{1.2}


```{r qc listing summary narrative, echo=FALSE, message=FALSE, warning=FALSE}
qc_summary_df <- qc_listing_df %>% group_by(WGS) %>% summarise(number = sum(!is.na(WGS)))
qc_summary_df <- qc_summary_df[order(-qc_summary_df$number),]

qc_tl <- nrow(qc_listing_df)
qc_distinct <- nrow(qc_summary_df)

qc_pass <- nrow(wgs_df[wgs_df$result == 'PASS', ])
qc_warning <- nrow(wgs_df[wgs_df$result == 'WARNING', ])

qc_pass_pc <- if (qc_pass != 0) (qc_pass/qc_tl) * 100 else 0
qc_pass_pc <- format(round(qc_pass_pc, 2), nsmall = 2)

qc_warning_pc <- if (qc_warning != 0) (qc_warning/qc_tl) * 100 else 0
qc_warning_pc <- format(round(qc_warning_pc, 2), nsmall = 2)

qc_concordance <- nrow(qc_listing_nc[qc_listing_nc$concordant == TRUE, ])
qc_cn_pc <- if (qc_concordance != 0) (qc_concordance/qc_tl) * 100 else 0
qc_cn_pc <- format(round(qc_cn_pc, 2), nsmall = 2)

```

$\\$
$\\$

```{r qc listing summary table, echo=FALSE, message=FALSE, warning=FALSE}
colnames(qc_summary_df) <- c('WGS_ID','Number')

qc_summary_df %>% 
  kable(booktabs = T, align = "lc")%>%
  column_spec(1, width = "8cm") %>%
  row_spec(0, background = "#D4D4D4",bold = TRUE)

```
* $\color{red}`r qc_distinct`$ distinct species were identified among $\color{red}`r qc_tl`$
isolates.

* $\color{red}`r qc_pass_pc`$ % (n=`r qc_pass`) of the isolates passed the QC, while $\color{red}`r qc_warning_pc`$ % (n=`r qc_warning`) were tagged with warning.


* Concordance between ARSRL and WGS species report was $\color{red}`r qc_cn_pc`$ %.


$\\$

### GRAPHS

\fontsize{7}{8}
\selectfont
\captionsetup[table]{labelformat=empty}
\renewcommand{\arraystretch}{1.2}



```{r pie_chart, echo=FALSE, message=FALSE, warning=FALSE}
# Calculate the counts for each "wgs_id"
wgs_id_count <- table(wgs_df$wgs_id)

# Create a data frame with "wgs_id" and their counts
wgs_count_df <- data.frame(wgs_id = names(wgs_id_count), Frequency = as.vector(wgs_id_count))

# Sort the data frame by the "Frequency" column in descending order
wgs_count_df <- wgs_count_df[order(-wgs_count_df$Frequency),]

# Now, set the factor levels in reverse order for correct sorting
wgs_count_df$wgs_id <- factor(wgs_count_df$wgs_id, levels = rev(wgs_count_df$wgs_id))


ggplot(data = wgs_count_df, aes(x = " ", y = Frequency, fill = wgs_id)) +
  geom_col(color = "black") +
  theme(legend.text = element_text(size=5)) +
  geom_text(aes(label = Frequency),
            position = position_stack(vjust = 0.5)) +
  xlab("") +
  labs(fill="WGS ID") +
  coord_polar(theta = "y") +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.grid  = element_blank())

```


### Result Classification
```{r organism results, echo=FALSE, message=FALSE, warning=FALSE}
# Define the desired order for the 'result' levels
desired_order <- c("PASS", "WARNING", "FAILURE")
wgs_df$result <- factor(wgs_df$result, levels = desired_order)

result_colors <- c("FAILURE" = "#f3533a", "PASS" = "#8ad879", "WARNING" = "#fa9f42")

ggplot(data = wgs_df, aes(x = ORG, fill = result)) +
  geom_bar() +
  facet_wrap(~ result) +
  coord_flip() +
  scale_fill_manual(values = result_colors) + # Apply the defined colors
  ylab("Frequency") +
  xlab("Organisms") +
  guides(fill = FALSE) #+  # Hide the legend for the 'result' variable
  #theme_minimal()
```


### Number of contigs
```{r ,echo=FALSE, message=FALSE, warning=FALSE}

ggplot(data = wgs_df) +
  geom_point(mapping = aes(x = no_of_contigs_value, y = ORG, color = ifelse(no_of_contigs_value < 301, "#8ad879", "#f3533a"))) +
  ylab("Organisms") +
  xlab("Number of Contigs") +
  scale_color_manual(values = c("#8ad879", "#f3533a"), labels = c("Passed", "Above threshold")) +
  labs(color = "Contigs")
  #theme_minimal()
```

### List of samples above/below QC threshold metrics
```{r QC failed list,echo=FALSE, message=FALSE, warning=FALSE}
wgs_df$sample_id <- gsub("\\_", "_", wgs_df$sample_id, fixed=TRUE)
QC_failed <- wgs_df %>%
  filter(result == "FAILURE") %>%
  select(sample_id,  result, contamination_result, no_of_contigs_value, n50_value, total_length_value)

if (nrow(QC_failed) > 0) {
  kable(QC_failed)
} else {
  cat("No QC failures found.")
}
```



### N50 Value
```{r n50_result ,echo=FALSE, message=FALSE, warning=FALSE}
n50_result_count <- wgs_df %>% 
  count(n50_result)



result_colors <- c("FAILURE" = "#f3533a", "PASS" = "#8ad879", "WARNING" = "#fa9f42")

ggplot(data = wgs_df) +
  geom_point(mapping = aes(x = n50_value, y = ORG, color = ifelse(n50_result == "PASS", "#8ad879", ifelse(n50_result == "WARNING", "#fa9f42", "#f3533a")))) +
  scale_color_manual(values = c("#8ad879", "#fa9f42","#f3533a"), labels = c("Passed", "Warning", "Failed")) +
  labs(color = "N50 Result") +
  xlab("N50 Value") +
  ylab("Organisms") +
  scale_x_continuous(
  labels = scales::comma_format(big.mark = ',',
                                 decimal.mark = '.'))


```



### Total Length 
```{r length_result ,echo=FALSE, message=FALSE, warning=FALSE}
length_result_count <- wgs_df %>% 
  count(total_length_result)



result_colors <- c("FAILURE" = "#f3533a", "PASS" = "#8ad879", "WARNING" = "#fa9f42")

ggplot(data = wgs_df) +
  geom_point(mapping = aes(x = total_length_value, y = ORG, color = ifelse(total_length_result == "PASS", "#8ad879", ifelse(total_length_result == "WARNING", "#fa9f42", "#f3533a")))) +
  scale_color_manual(values = c("#8ad879", "#fa9f42","#f3533a"), labels = c("Passed", "Warning", "Failed")) +
  labs(color = "Total length result") +
  xlab("Total length Value") +
  ylab("Organisms") +
  scale_x_continuous(
  labels = scales::comma_format(big.mark = ',',
                                 decimal.mark = '.'))


```

