---
title: \vspace{-1.5cm} \begin{LARGE} WGS Quality Control Report \end{LARGE}
output:
  pdf_document: 
    latex_engine: lualatex
    keep_tex: true
geometry: margin=0.75in
mainfont: Helvetica
papersize: a4
header-includes:
- \usepackage{titling}
- \pretitle{\begin{flushleft}}
- \posttitle{\end{flushleft}} 
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{float}
- \floatplacement{figure}{H}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{makecell}
- \usepackage{xcolor}
- \usepackage[table]{xcolor}
- \usepackage{soul}
- \usepackage{caption}
- \usepackage[singlelinecheck=false]{caption}
- \usepackage[font={small,bf}]{caption}
- \usepackage{multirow}
- \usepackage{array}
- \usepackage{lscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage[dvipsnames]{xcolor}
- \renewcommand{\footnotesize}{\tiny}
- \usepackage{threeparttable}
- \definecolor{mygreen}{HTML}{006b54}  #define new hex 
- \definecolor{myyellow}{HTML}{ffd858}
- \definecolor{myred}{HTML}{D2042D}

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.pos = 'H')
```

```{r load library, echo = FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(knitr)
library(readxl)
library(dplyr)
library(kableExtra)
library(RColorBrewer)
library(scales)
library(readr)
library(writexl)
library(conflicted)
library(stringr)
library(ggplot2)
library(plotly)

conflicts_prefer(openxlsx::write.xlsx)
conflicts_prefer(dplyr::filter)
```




```{r load-data, echo = FALSE, warning = FALSE, message = FALSE}
# --- rename species column to wgs_org ---
wgs_df <- wgs_result_df %>% rename(wgs_id = species)

# --- complete wgs_org  ---
wgs_df$wgs_id <- ifelse(wgs_df$wgs_id == "Acinetobacter", "Acinetobacter baumannii", 
                         ifelse(wgs_df$wgs_id == "Escherichia", "Escherichia coli", wgs_df$wgs_id))

# --- load organism code reference file ---
org_df <- read_xlsx("data_files/whonet_codes_2024.xlsx", "ORGANISM") %>%
  select(ORG_ARS, ORGANISM_name) %>% 
  rename(wgs_id = ORGANISM_name)

# --- merge dataframes ---
wgs_df <- merge(wgs_df, org_df, by = "wgs_id", all.x = TRUE)


# --- rename ORG_ARS column to ORG ---
wgs_df <- wgs_df %>% rename(org_code = ORG_ARS)

# --- Clean organism name ---
wgs_df <- wgs_df %>%
  mutate(
    org_code_clean = case_when(
      wgs_id == "Escherichia/Shigella" ~ "eco",
      str_detect(wgs_id, "Salmonella enterica") ~ "sat",
      str_detect(wgs_id, "Salmonella") ~ "sal",
      str_detect(wgs_id, "Streptococcus pseudopneumoniae") ~ "spn",
      str_detect(wgs_id, "Streptococcus pyogenes") ~ "spy",
      str_detect(wgs_id, "Streptococcus oralis") ~ "sol",
      str_detect(wgs_id, "Streptococcus") ~ "spn",
      str_detect(wgs_id, "Staphylococcus epidermidis") ~ "sep",
      str_detect(wgs_id, "Staphylococcus") ~ "sau",
      str_detect(wgs_id, "Shigella") ~ "shi",
      str_detect(wgs_id, "Vibrio") ~ "vic",
      str_detect(wgs_id, "Klebsiella") ~ "kpn",
      str_detect(wgs_id, "Raoultella planticola") ~ "kpn",
      str_detect(wgs_id, "Pseudomonas") ~ "pae",
      str_detect(wgs_id, "Acinetobacter") ~ "aba",
      str_detect(wgs_id, "Burkholderia") ~ "pce",
      str_detect(wgs_id, "Serratia marcescens") ~ "sma",
      TRUE ~ org_code
    )
  )

```




```{r sample-sheet, echo = FALSE, warning = FALSE, message = FALSE}
# --- Read samplesheet ---
wgs_samplesheet <- read_csv(paste0("data_files/", get_samplesheet, ".csv")) 

# --- Get experiment name ---
WGS_exp_name <- unique(wgs_samplesheet$`experiment name`)

# --- remove special characters in batch name ---
batch_name_clean <- gsub("[[:punct:]]", "", batch_code)

# --- Modify dataframe ---
wgs_samplesheet <- wgs_samplesheet%>%
  rename_with(tolower) %>%              # lowercase column names
  mutate(
    sample_id = toupper(sample_id),     # uppercase sample_id
    sample_id = str_replace_all(sample_id, "-", "_"),  # replace "-" with "_"
    batch_name_clean = str_remove_all(batch_code, "[[:punct:]]"), # clean batch code
    sample_id = paste(sample_id, batch_name_clean, sep = "_")      # append batch
  ) %>%
  select(sample_id, description)  # optional: remove helper column

```


\normalsize Batch Name: `r batch_code`

\normalsize Experiment Name: `r WGS_exp_name`

\fontsize{7}{8}
\selectfont
\captionsetup[table]{labelformat=empty}
\renewcommand{\arraystretch}{1.2}
```{r wgs-data, echo = FALSE, warning = FALSE, message = FALSE}
# --- Subset wgs_df ---
wgs_id_df <- subset(wgs_df, select = c('sample_id','wgs_id'))


# --- paste batch number to arsrl result sample_id ---
arsrl_id_df <- arsrl_result_df %>%
  mutate(sample_id = paste(sample_id, batch_name_clean, sep = "_"))


# --- Add isolate numbers ---
arsrl_id_df$iso_num <- seq_len(nrow(arsrl_id_df))


# --- merge arsrl result  and samplesheet ---
arsrl_result_merge <- merge(unique(arsrl_id_df), unique(wgs_samplesheet), by = "sample_id", all.x= TRUE)


# --- Merge WGS and ARSRL results ---
qc_listing_df <- merge(
  unique(wgs_id_df), 
  unique(arsrl_result_merge), 
  by = "sample_id", 
  all.x = TRUE
)

# Preserve original WGS order
qc_listing_df <- qc_listing_df[match(wgs_result_df$sample_id, qc_listing_df$sample_id), ]

# Fill missing WGS IDs
qc_listing_df$wgs_id[is.na(qc_listing_df$wgs_id)] <- "Not Identified"

# --- Handle samples with no WGS results ---
wgs_no_result <- wgs_samplesheet[!wgs_samplesheet$sample_id %in% wgs_df$sample_id, ]

if (nrow(wgs_no_result) > 0) {
  qc_listing_df$description <- ifelse(
    is.na(qc_listing_df$description),
    wgs_samplesheet_subset$description[match(
      paste0(wgs_no_result$sample_id, "_", batch_name_clean),
      wgs_samplesheet_subset$sample_id
    )],
    qc_listing_df$description
  )
}


# --- Concordance check ---
qc_listing_df <- qc_listing_df %>%
  rowwise() %>%
  mutate(
    Genus = strsplit(as.character(arsrl_org), " ")[[1]][1],
    concordant = ifelse(wgs_id != "Not Identified", grepl(Genus, wgs_id), TRUE),
    concordant = ifelse(arsrl_org == "No Referred Data", TRUE, concordant)
  ) %>%
  ungroup()

# Append (x) to non-concordant ARSRL results
qc_listing_df$arsrl_org <- ifelse(
  qc_listing_df$concordant,
  qc_listing_df$arsrl_org,
  paste0(qc_listing_df$arsrl_org, " (x)")
)

# Match concordance back to WGS df
wgs_df$concordant <- qc_listing_df$concordant[match(wgs_df$sample_id, qc_listing_df$sample_id)]


# --- Modify and Filter Bactscout dataframe ---
df_bactscout_mod <- df_bactscout %>%
  mutate(name = gsub("-", "_", name)) %>%
  filter(name %in% sample_name)





```


```{r sunburst-graph, echo = FALSE, warning = FALSE, message = FALSE, fig.height=10, fig.width=9}
final_result_df <- df_bactscout_mod %>%
  select(species, a_final_status) 

sunburst_df <- final_result_df %>%
  count(species, a_final_status, name = "value") %>%
  mutate(
    label = a_final_status,
    parent = species
  )

# Add species as root nodes
species_df <- sunburst_df %>%
  group_by(species) %>%
  summarise(value = sum(value), .groups = "drop") %>%
  mutate(
    label = species,
    parent = ""
  )

sunburst_data <- bind_rows(
  species_df %>% select(label, parent, value),
  sunburst_df %>% select(label, parent, value)
)


sunburst_data <- sunburst_data %>%
  mutate(
    bold_label = paste0("<b>", label, "</b>"),
    ids = ifelse(parent == "", label, paste(parent, "-", label)),
    color = case_when(
      label == "PASSED"  ~ "#006b54",
      label == "FAILED"  ~ "#D2042D",
      label == "WARNING" ~ "#ffd858",
      TRUE               ~ "#aec7e8" # Default for bacteria
    )
  )


# 3. Create the plot
# 3. Create the plot with bold font settings
fig <- plot_ly(
  data = sunburst_data,
  ids = ~ids,
  labels = ~bold_label,
  parents = ~parent,
  values = ~value,
  type = 'sunburst',
  branchvalues = 'total',
  marker = list(colors = ~color),
  # Explicitly setting font to bold
  insidetextfont = list(size = 15),
  outsidetextfont = list(size = 15),
  hovertemplate = "<b>%{label}</b><br>Value: %{value}<extra></extra>"
)

fig <- fig %>% plotly::layout(
  title = list(
    text = "<b><i>Figure 1.</i></b> <b>Distribution and quality control results of sequenced isolates.</b>",
    font = list(size = 12),
    y = 0.05,       # Lift it slightly off the absolute edge
    x = 0.5,
    xanchor = 'center',
    yanchor = 'bottom' # Anchor the bottom of the text
  ),
  # Increase 'b' (bottom) margin to at least 50-80 to fit the text
  margin = list(l = 0, r = 0, b = 80, t = 40) 
)

fig



```


\small\textbf{\emph{Table 1.} Concordance analysis of bacterial identification of sequenced isolates.}

```{r qc-listing, echo = FALSE, warning = FALSE, message = FALSE}
# --- Identify failed WGS sample names ---
raw_read_failed <- df_bactscout_mod$name[df_bactscout_mod$a_final_status == "FAILED"]

  
# --- Get row numbers in qc_listing_df where sample_id matches any failed sample ---
failure_color <- which(qc_listing_df$sample_id %in% raw_read_failed)


# --- Clean up ---
qc_listing_df <- qc_listing_df %>%
  select(-Genus, -concordant) %>%
  mutate(sample_id = sub("(.*)_[^_]*$", "\\1", sample_id))


# Rename and reorder columns
colnames(qc_listing_df) <- c("Sample ID", "WGS", "ARSRL", "Isolate No.", "Description")
qc_listing_df <- qc_listing_df[, c("Isolate No.", "Sample ID", "Description", "ARSRL", "WGS")]

# --- Generate formatted table ---
qc_listing_df %>%
  kable(booktabs = TRUE, align = "c", escape = FALSE, row.names = FALSE) %>%
  column_spec(1, width = "1cm") %>%
  column_spec(2, width = "2.8cm") %>%
  column_spec(3, width = "1.5cm") %>%
  column_spec(4, width = "5cm", italic = TRUE) %>%
  column_spec(5, width = "5cm", italic = TRUE) %>%
  row_spec(failure_color, background = "#FD7979") %>%
  row_spec(0, background = "#D4D4D4", bold = TRUE, align = "c") %>%
  footnote(
    general = c(
      "PASS", "  |  ", 
      "\\\\colorbox{Salmon}{FAILURE}", "  |  ",
      "(x) - NON-CONCORDANT", "  |"
    ), 
    general_title = "Legend:",
    footnote_as_chunk = TRUE, 
    escape = FALSE
  )


```



```{r low-reads, echo=FALSE, message=FALSE, warning=FALSE,results='asis'}
#filter dataframe based on the sample_id not present in the WGS samplesheet
wgs_lowreads <- wgs_samplesheet[!(wgs_samplesheet$sample_id %in% wgs_df$sample_id), ]
wgs_lowreads <- subset(wgs_lowreads, select = c('sample_id','description'))

if (nrow(wgs_lowreads != 0)){
  lowreads_table <- wgs_lowreads
  lowreads_table$remarks <- c("low read count")
  #remove batch name in sample_id
  lowreads_table$sample_id <- sub("(.*)_[^_]*$", "\\1", lowreads_table$sample_id)
  colnames(lowreads_table) <- c('Sample ID','Description','Remarks') 
  
  kable(lowreads_table,  "latex", booktabs = T, align = "c") %>%
    kable_styling(position = "left") %>%
    column_spec(1, width = "3cm") %>%
    column_spec(2, width = "3cm") %>%
    column_spec(3, width = "7cm") %>%
    add_header_above(c("Sample excluded in the analysis" = 3), align = "l", bold = TRUE) %>%
    row_spec(0, background = "#D4D4D4",bold = TRUE,align = "c")
}

```





$\\$
\newpage
\begin{landscape}
\fontsize{7}{8}
\selectfont
\captionsetup[table]{labelformat=empty}
\renewcommand{\arraystretch}{1.2}
\section*{RAW READS QC}

\small\textbf{\emph{Table 2.} Summary of raw reads quality control analysis.}



```{r bactscout-df-value, echo = FALSE, warning = FALSE, message = FALSE}
# --- Create new column to check final status with warning status and list column indices with failed results ---
bactscout_df_result <- df_bactscout_mod %>%
  select(name, species, Contamination = contamination_message,`Coverage (Qualibact)` = coverage_estimate_qualibact_status, `Duplication Rate` = duplication_status, 
        `GC Content` = gc_content_status, `MLST (ST)` = mlst_status, `Read Length` = read_length_status, `Read Q30` = read_q30_status, a_final_status) %>%
  mutate(Contamination = ifelse(Contamination == "OK", "PASSED", "FAILED"))


bactscout_df_result$failed_idx <- apply(bactscout_df_result, 1, function(row) {
  if (row["a_final_status"] == "FAILED") {
    # Get indices of columns with FAILED
    failed_cols <- which(row[-c(1,2,10)] == "FAILED") # exclude name and species
    # Adjust index to match original dataframe
    failed_cols <- failed_cols + 3
    paste(failed_cols, collapse = ", ")
  } else {
    "" 
  }
})


bactscout_df_result$failed_cols <- apply(bactscout_df_result, 1, function(row) {
  if (row["a_final_status"] == "FAILED") {
    warning_cols <- names(row)[which(row[-c(1,2,10)] == "FAILED") + 2] # adjust for original df
    if (length(warning_cols) == 0) {
      ""
    } else {
      paste(warning_cols, collapse = ", ")
    }

  } else {
    "" 
  }
})



# --- List sample without bactscout result ---
no_bactscout_result <- setdiff(sample_name, df_bactscout_mod$name)

# --- remove text after last underscore ---
no_bactscout_result <- sub("_[^_]*$", "", no_bactscout_result)


# --- Modify dataframe ---
bactscout_df_value <- df_bactscout_mod %>%
  select(name, species, species_abundance, coverage_estimate_qualibact, duplication_rate, 
         gc_content, mlst_st, read1_mean_length, read2_mean_length, read_q30_rate, a_final_status) %>%
  rename(sample_id = name) %>%
  mutate(read_length = paste0(read1_mean_length, " - ", read2_mean_length)) %>%
  select(-read1_mean_length, -read2_mean_length) %>%
  relocate(read_length, .before = "read_q30_rate")


# Change MLST ST column NUll value to Blank
bactscout_df_value$mlst_st <- lapply(bactscout_df_value$mlst_st, function(x) {
  if (is.null(x)) "" else x
})



# --- Get Isolate number by merging ---
bactscout_df_merge <- merge(bactscout_df_value, arsrl_id_df, by ="sample_id")

# --- Modify dataframe ---
bactscout_df_merge <- bactscout_df_merge %>%
  mutate(sample_id = sub("(.*)_[^_]*$", "\\1", sample_id)) %>%
  select(-arsrl_org) %>%
  relocate(iso_num, .before = "sample_id") 

bactscout_df_merge$sample_id <- gsub("_", "\\_", bactscout_df_merge$sample_id, fixed = TRUE)

# Rename and reorder columns
colnames(bactscout_df_merge) <- c("Isolate No.","Sample ID", "Species", "Species Abundance", "Coverage (Qualibact)", "Duplication Rate",
                                 "GC Content", "MLST (ST)", "Read Length", "Read Q30", "Final Status")


for (i in 1:nrow(bactscout_df_result)) {
  if (!is.na(bactscout_df_result$failed_idx[i]) && nzchar(bactscout_df_result$failed_idx[i])) {
    # Split and convert to numeric safely
      indices <- tryCatch({
        as.numeric(trimws(unlist(strsplit(bactscout_df_result$failed_idx[i], ","))))
      }, error = function(e) numeric(0))
      
      # Remove any NAs from conversion
      indices <- indices[!is.na(indices) & indices > 0]
      
      # Apply formatting to each specified column
      for (col in indices) {
        # Adjust column index (now accounts for renamed columns)
        display_col <- col
        #display_col <- col + 3  # Adjust based on your column structure
        
        if (display_col <= ncol(bactscout_df_merge)) {
          # Get original value
          original_value <- bactscout_df_merge[i, display_col]  # Adjust offset as needed
          
          # Apply LaTeX formatting for PDF
          bactscout_df_merge[i, display_col] <- paste0("\\colorbox{myred}{\\textcolor{white}{\\textbf{",as.character(original_value),"}}}")

        }
        
      }
      
  }
}



# Get row indices
rows_passed <- which(bactscout_df_result$a_final_status == 'PASSED')
rows_warning <- which(bactscout_df_result$a_final_status == 'WARNING')
rows_failed <- which(bactscout_df_result$a_final_status == 'FAILED')


          
bactscout_df_merge <- bactscout_df_merge %>%
  mutate(
    `Final Status` = case_when(
      row_number() %in% rows_passed  ~ cell_spec(`Final Status`, background = "#006b54", color = "white"),
      row_number() %in% rows_warning ~ cell_spec(`Final Status`, background = "#ffd858"),
      row_number() %in% rows_failed  ~ cell_spec(`Final Status`, background = "#D2042D", color = "white"),
      TRUE ~ `Final Status`
      )
    )


# --- Generate formatted table ---
# Create the table
bactscout_df_merge %>%  
  kable(format = "latex", booktabs = TRUE, escape = FALSE, align = "c", row.names = FALSE) %>%
  kable_styling(latex_options = c("scale_down", "hold_position")) %>%
  row_spec(0, background = "#D4D4D4",bold = TRUE) %>%
  column_spec(3,italic = TRUE) %>%
  footnote(general = c("\\\\colorbox{mygreen}{\\\\textcolor{white}{PASSED}}","  |  ", 
                       "\\\\colorbox{myyellow}{WARNING}","  |  ",  
                      "\\\\colorbox{myred}{\\\\textcolor{white}{FAILED}}"), 
          general_title = "Legend:",
          footnote_as_chunk = T, escape = F)

  
```





```{r no-bactscout-result ,echo=FALSE, message=FALSE, warning=FALSE}
if(length(no_bactscout_result) != 0){
  # Convert the vector to a single comma-separated string
  # The collapse argument specifies the string to be inserted between each element
  no_bactscout_id <- paste(no_bactscout_result, collapse = ", ")
  
  # Create new dataframe
  df_no_bactscout <- data.frame(sample_id = no_bactscout_id, Remarks = "No Raw Reads Data")
  
  colnames(df_no_bactscout) <- c('Sample ID','Remarks')
  
  no_bactscout_table <- df_no_bactscout %>% 
    kable(format = "latex", booktabs = TRUE, escape = TRUE, align = c("l", "c")) %>%
    kable_styling(
      position = "left",
      latex_options = c("hold_position")
    ) %>%
    column_spec(1, width = "20cm") %>%
    column_spec(2, width = "5cm") %>%
    row_spec(0, background = "#D4D4D4", bold = TRUE) %>%
    add_header_above(c("Summary of No Raw Reads Data" = 2), align = "l", bold = TRUE)
  
    no_bactscout_table  
  
  
}


```


$\\$ 


```{r recommendation-table ,echo=FALSE, message=FALSE, warning=FALSE}
reco_df <- bind_rows(
  # Non-concordant WGS
  wgs_df %>%
    filter(concordant == "FALSE") %>%
    transmute(sample_id, failed_conditions = "Non-concordant result"),

  # Failed raw reads (from BactScout)
  bactscout_df_result %>%
    filter(failed_cols != "") %>%
    transmute(sample_id = name,
              failed_conditions = failed_cols),

  # Low read counts (only if exists)
  if (nrow(wgs_lowreads) > 0)
    wgs_lowreads %>%
      transmute(sample_id,
                failed_conditions = "Low read counts")
) %>%
  mutate(sample_id = sub("(.*)_[^_]*$", "\\1", sample_id)) %>%   # 👈 normalize IDs
  group_by(sample_id) %>%
  summarise(
    failed_conditions = paste(unique(failed_conditions), collapse = ", "),
    .groups = "drop"
  )



if (nrow(reco_df) == 0) { failed_conditions <- "No further action required for this batch."}

colnames(reco_df) <- c('Sample ID','Reason - Failed Metrics')
reco_table <- reco_df %>% 
  kable(
    format = "latex",
    booktabs = TRUE,
    escape = TRUE,  # Keep FALSE since you pre-escaped underscores
    align = c("l", "c")
  ) %>%
  kable_styling(
    position = "left",
    latex_options = c("striped", "hold_position")
  ) %>%
  column_spec(1, width = "6cm") %>%
  column_spec(2, width = "6cm") %>%
  row_spec(0, background = "#D4D4D4", bold = TRUE) %>%
  add_header_above(c("Summary of Raw Reads with FAILED QC results" = 2), align = "l", bold = TRUE)

  reco_table  


```


\end{landscape}




\begin{landscape}
\fontsize{7}{8}
\selectfont
\captionsetup[table]{labelformat=empty}
\renewcommand{\arraystretch}{1.2}
\section*{ASSEMBLY QC}

\small\textbf{\emph{Table 3.} Summary of assembly reads quality control analysis.}



```{r assembly-qc, echo = FALSE, warning = FALSE, message = FALSE}
assembly_qc <- merge(wgs_df, arsrl_id_df, by = "sample_id")


# Check if wgs_id is genus only
assembly_qc$with_species<- str_detect(assembly_qc$wgs_id, " ") 


# Filter df with genus only
genus_df <- assembly_qc %>%
  filter(with_species == FALSE)


# --- Read samplesheet ---
qualibact_data <- read_csv("data_files/qualibact_filtered_metrics.csv") %>%
  mutate(species = gsub("_", " ", species))


if (nrow(genus_df) != 0) {
  # Create the regex pattern
  wgs_pattern <- paste(genus_df$wgs_id, collapse = "|")
  
  qualibact_partial <- qualibact_data %>%
    # Use the column name directly
    filter(grepl(wgs_pattern, species)) %>% 
    mutate(metric_clean = str_to_lower(metric))
  
  result_qualibact <- qualibact_partial %>%
    mutate(species = sub(" .*", "", species)) %>%
    group_by(species, metric_clean) %>%
    summarize(
      lower_bounds = min(lower_bounds, na.rm = TRUE),
      upper_bounds = max(upper_bounds, na.rm = TRUE),
      .groups = "drop"
    )
  
  # --- Prepare reference table: only species present in assembly_qc ---
  qualibact_ref <- qualibact_data %>%
    filter(species %in% assembly_qc$wgs_id) %>%
    mutate(metric_clean = str_to_lower(metric)) %>%
    select(-metric) %>%
    relocate(metric_clean, .after = "species")
  
  
  qualibact_ref <- rbind(qualibact_ref, result_qualibact)
  
  # -- Replace -Inf in dataframe with NA --
  qualibact_ref[sapply(qualibact_ref, is.infinite)] <- NA
  
} else {
  # --- Prepare reference table: only species present in assembly_qc ---
  qualibact_ref <- qualibact_data %>%
    filter(species %in% assembly_qc$wgs_id) %>%
    mutate(metric_clean = str_to_lower(metric))

}



# --- Map qualibact metrics to assembly_qc column names ---
metric_map <- c(
  "completeness"             = "completeness",
  "contamination"            = "contamination",
  "total_coding_sequences"   = "total_coding_sequences",
  "genome_size"              = "genome_size",
  "gc_content"               = "gc_content",
  "n50"                      = "n50_contig_length",
  "no_of_contigs"            = "total_contig"
)

# --- QC check function ---
check_bounds <- function(value, lower, upper) {
  case_when(
    !is.na(lower) & value < lower ~ "FAILED",
    !is.na(upper) & value > upper ~ "FAILED",
    TRUE                          ~ "PASSED"
  )
}

# --- Apply QC checks ---
assembly_qc_checked <- assembly_qc

for (m in names(metric_map)) {

  ref <- qualibact_ref %>%
    filter(metric_clean == m) %>%
    select(species, lower_bounds, upper_bounds)

  assembly_qc_checked <- assembly_qc_checked %>%
    left_join(ref, by = c("wgs_id" = "species")) %>%
    mutate(
      !!paste0(metric_map[m], "_qc") :=
        check_bounds(.data[[metric_map[m]]], lower_bounds, upper_bounds)
    ) %>%
    select(-lower_bounds, -upper_bounds)
}



# --- List all FAILED QC ---
assembly_qc_checked <- assembly_qc_checked %>%
  rowwise() %>%
  mutate(
    condition_checks = list(c(
      "Completeness" = completeness_qc == "PASSED",
      "Contamination" = contamination_qc == "PASSED",
      "Total Coding Sequences" = total_coding_sequences_qc == "PASSED",
      "Genome Size" = genome_size_qc == "PASSED",
      "GC Content" = gc_content_qc == "PASSED",
      "N50" = n50_contig_length_qc == "PASSED",
      "Total Contigs" = total_contig_qc == "PASSED"
    )),
    # Set result to FALSE if any condition is FALSE or NA
    result = all(unlist(condition_checks), na.rm = FALSE),
    
    # Only list failed conditions (FALSE), remove NA
    failed_conditions = {
      cond_values <- unlist(condition_checks)
      failed <- names(cond_values[!is.na(cond_values) & cond_values == FALSE])
      if (length(failed) == 0) NA_character_ else paste(failed, collapse = ", ")
    }
  ) %>%
  ungroup()



# Get raw reads result from bactscout_df_result
assembly_qc_checked$raw_reads_result <- bactscout_df_result$a_final_status[bactscout_df_result$name == assembly_qc_checked$sample_id]

# condition assembly qc result
assembly_qc_checked <- assembly_qc_checked %>%
  mutate(assembly_result = ifelse(result == TRUE, "PASSED", "FAILED"))

file_name <- paste0("data_files/assembly_qc_checked.xlsx")
write_xlsx(assembly_qc_checked, file_name)

```


```{r aasembly-qc-graph, echo = FALSE, warning = FALSE, message = FALSE}
# --- Create new column to check final status with warning status and list column indices with failed results ---
assembly_qc_result <- assembly_qc_checked %>%
  select(iso_num, sample_id, wgs_id, completeness_qc, contamination_qc, total_coding_sequences_qc, 
        genome_size_qc, gc_content_qc, n50_contig_length_qc, total_contig_qc)

assembly_qc_result$failed_idx <- apply(assembly_qc_result, 1, function(row) {
  failed_cols <- which(row[-c(1:3)] == "FAILED") # exclude name and species
    # Adjust index to match original dataframe
    failed_cols <- failed_cols + 3
    paste(failed_cols, collapse = ", ")
})


# --- List sample without assembly result ---
no_assembly_result <- setdiff(sample_name, assembly_qc_result$sample_id)

# --- remove text after last underscore ---
no_assembly_result <- sub("_[^_]*$", "", no_assembly_result)


# --- Modify dataframe ---
assembly_qc_value <- assembly_qc_checked %>%
  select(iso_num, sample_id, wgs_id, completeness, contamination, total_coding_sequences, 
        genome_size, gc_content, n50_contig_length, total_contig, raw_reads_result, assembly_result) %>%
  mutate(sample_id = sub("(.*)_[^_]*$", "\\1", sample_id))


assembly_qc_summary <- assembly_qc_value


# Create a conversion function to convert Basepairs to MEgabasepairs
bp_to_mb <- function(bp) {
  x <- bp / 1e6
  x <- format(round(x, 2), nsmall = 2)
  x <- paste0(x, " Mb")
  return(x)
}

assembly_qc_value$genome_size <- bp_to_mb(assembly_qc_value$genome_size)


assembly_qc_value$sample_id <- gsub("_", "\\_", assembly_qc_value$sample_id, fixed = TRUE)




# Rename and reorder columns
colnames(assembly_qc_value) <- c("Isolate No.","Sample ID", "WGS ID", "Completeness", "Contamination", "Total Coding Sequences",
                                 "Genome Size", "GC Content", "N50", "Total Contig", "Raw Reads Result", "Assembly Result")


for (i in 1:nrow(assembly_qc_result)) {
  if (!is.na(assembly_qc_result$failed_idx[i]) && nzchar(assembly_qc_result$failed_idx[i])) {
    # Split and convert to numeric safely
      indices <- tryCatch({
        as.numeric(trimws(unlist(strsplit(assembly_qc_result$failed_idx[i], ","))))
      }, error = function(e) numeric(0))
      
      # Remove any NAs from conversion
      indices <- indices[!is.na(indices) & indices > 0]
      
      # Apply formatting to each specified column
      for (col in indices) {
        # Adjust column index (now accounts for renamed columns)
        display_col <- col
        #display_col <- col + 3  # Adjust based on your column structure
        
        if (display_col <= ncol(assembly_qc_value)) {
          # Convert all columns to character (or only numeric ones)
          assembly_qc_value[] <- lapply(assembly_qc_value, as.character)
          
          # Get original value
          original_value <- assembly_qc_value[i, display_col]  # Adjust offset as needed
          
          # Apply LaTeX formatting for PDF
          assembly_qc_value[i, display_col] <- as.character(paste0("\\colorbox{myred}{\\textcolor{white}{\\textbf{", original_value, "}}}")
)

        }
        
      }
      
  }
}


# Change Assembly Result based on Raw Reads Result
assembly_qc_value <- assembly_qc_value %>%
  mutate(`Assembly Result` = ifelse(`Raw Reads Result` == 'FAILED', "FAILED", `Assembly Result`))


# Get row indices
rows_passed_raw <- which(assembly_qc_value$`Raw Reads Result`== 'PASSED')
rows_warning_raw <- which(assembly_qc_value$`Raw Reads Result` == 'WARNING')
rows_failed_raw <- which(assembly_qc_value$`Raw Reads Result` == 'FAILED')

rows_passed_asm <- which(assembly_qc_value$`Assembly Result`== 'PASSED')
rows_warning_asm <- which(assembly_qc_value$`Assembly Result` == 'WARNING')
rows_failed_asm <- which(assembly_qc_value$`Assembly Result` == 'FAILED')


          
assembly_qc_value <- assembly_qc_value %>%
  mutate(
    `Raw Reads Result` = case_when(
      row_number() %in% rows_passed_raw   ~ cell_spec(`Raw Reads Result`, background = "#006b54", color = "white"),
      row_number() %in% rows_warning_raw  ~ cell_spec(`Raw Reads Result`, background = "#ffd858"),
      row_number() %in% rows_failed_raw   ~ cell_spec(`Raw Reads Result`, background = "#D2042D", color = "white"),
      TRUE ~ `Raw Reads Result`
      )
    ) %>%
  mutate(
    `Assembly Result` = case_when(
      row_number() %in% rows_passed_asm   ~ cell_spec(`Assembly Result`, background = "#006b54", color = "white"),
      row_number() %in% rows_warning_asm  ~ cell_spec(`Assembly Result`, background = "#ffd858"),
      row_number() %in% rows_failed_asm   ~ cell_spec(`Assembly Result`, background = "#D2042D", color = "white"),
      TRUE ~ `Assembly Result`
      )
    )




# --- Generate formatted table ---
# Create the table
assembly_qc_value %>%  
  kable(format = "latex", booktabs = TRUE, escape = FALSE, align = "c", row.names = FALSE) %>%
  kable_styling(latex_options = c("scale_down", "hold_position")) %>%
  row_spec(0, background = "#D4D4D4",bold = TRUE) %>%
  column_spec(3,italic = TRUE) %>%
  footnote(general = c("\\\\colorbox{mygreen}{\\\\textcolor{white}{PASSED}}","  |  ", 
                       "\\\\colorbox{myyellow}{WARNING}","  |  ",  
                      "\\\\colorbox{myred}{\\\\textcolor{white}{FAILED}}"), 
          general_title = "Legend:",
          footnote_as_chunk = T, escape = F)

  
```


```{r assembly-recommendation-table ,echo=FALSE, message=FALSE, warning=FALSE}
assembly_failed <- assembly_qc_checked %>% filter(!is.na(failed_conditions))

if (nrow(assembly_failed) !=0) {
  assembly_reco_df <- assembly_failed %>% select(sample_id, failed_conditions) %>%
    mutate(sample_id = sub("(.*)_[^_]*$", "\\1", sample_id))
} else {
  sample_id <- ""
  failed_conditions <- "No further action required for this batch."

  
  assembly_reco_df <- data.frame(sample_id,failed_conditions, remarks)
}
  

colnames(assembly_reco_df) <- c('Sample ID','Reason - Failed Metrics')
assembly_reco_table <- assembly_reco_df %>% 
  kable(
    format = "latex",
    booktabs = TRUE,
    escape = TRUE,  # Keep FALSE since you pre-escaped underscores
    align = c("l", "c", "c")
  ) %>%
  kable_styling(
    position = "left",
    latex_options = c("striped", "hold_position")
  ) %>%
  column_spec(1, width = "6cm") %>%
  column_spec(2, width = "10cm") %>%
  row_spec(0, background = "#D4D4D4", bold = TRUE) %>%
  add_header_above(c("Summary of Assembly Reads with FAILED QC results" = 2), align = "l", bold = TRUE)

  assembly_reco_table  


```

\section*{Final Recommendation based on QC results}



```{r overall-recommendation-table ,echo=FALSE, message=FALSE, warning=FALSE}
overall_reco <- assembly_qc_checked %>% 
  filter(raw_reads_result == "FAILED" | assembly_result == "FAILED") %>%
  select(sample_id, raw_reads_result, assembly_result) %>%
  mutate(sample_id = sub("(.*)_[^_]*$", "\\1", sample_id)) %>%
  group_by(raw_reads_result, assembly_result) %>%
  summarize(sample_id = paste0(sample_id, collapse = ", "))

# Generate Final Result
overall_reco <- overall_reco %>%
  mutate(final_result = case_when(
    raw_reads_result == "PASSED" & assembly_result == "PASSED" ~ "PASSED",
    raw_reads_result == "FAILED" | assembly_result == "FAILED" ~ "FAILED",
    .default = ""
  ))


if (nrow(overall_reco) !=0) {
  overall_reco_df <- overall_reco %>%
    relocate(sample_id, .before = raw_reads_result)
  
  colnames(overall_reco_df) <- c("Sample ID", "Raw Reads Result", "Assembly Result", "Final Result")
  
  overall_reco_table <- overall_reco_df %>% 
  kable(
    format = "latex",
    booktabs = TRUE,
    escape = TRUE,  # Keep FALSE since you pre-escaped underscores
    align = c("l", "c", "c")
  ) %>%
  kable_styling(
    position = "left",
    latex_options = c("striped", "hold_position")
  ) %>%
  column_spec(1, width = "8cm") %>%
  column_spec(2:4, width = "3cm") %>%
  row_spec(0, background = "#D4D4D4", bold = TRUE)

  overall_reco_table  
  
  
  
} else {
  sample_id <- ""
  remarks <- "No further action required for this batch."

  overall_reco_df <- data.frame(sample_id,remarks)
  colnames(overall_reco_df) <- c("Sample ID", "Remarks")
  
  overall_reco_table <- overall_reco_df %>% 
  kable(
    format = "latex",
    booktabs = TRUE,
    escape = TRUE,  # Keep FALSE since you pre-escaped underscores
    align = c("l", "c", "c")
  ) %>%
  kable_styling(
    position = "left",
    latex_options = c("striped", "hold_position")
  ) %>%
  column_spec(1, width = "6cm") %>%
  column_spec(2, width = "6cm") %>%
  row_spec(0, background = "#D4D4D4", bold = TRUE)

  overall_reco_table  
}
  




```



```{r overall-result-summary, echo = FALSE, warning = FALSE, message = FALSE}
raw_reads_summary <- bactscout_df_value %>%
  select(-a_final_status, -species) 

# Get raw reads failed result from bactscout_df_result
raw_reads_summary$failed_conditions <- bactscout_df_result$failed_cols[bactscout_df_result$name == raw_reads_summary$sample_id]

# create new wgs_summary_table with additional column for result
raw_reads_summary <- raw_reads_summary %>%
  mutate(
    batch_code = batch_code,
    sample_id = sub("(.*)_[^_]*$", "\\1", sample_id)
    ) %>%
  relocate(batch_code, .before = sample_id) %>%
  rename(raw_failed_conditions = failed_conditions)


overall_result <- merge(raw_reads_summary, assembly_qc_summary, by ="sample_id")


# --- Get assembly result failed condition --
failed_assembly <- assembly_qc_checked %>%
  select(sample_id, failed_conditions) %>%
  mutate(sample_id = sub("(.*)_[^_]*$", "\\1", sample_id)) %>%
  rename(assembly_failed_conditions = failed_conditions)


# -- Merge --
overall_result <- merge(overall_result, failed_assembly, by ="sample_id")


overall_result <- overall_result %>%
  mutate(final_result = case_when(
    raw_reads_result == "PASSED" & assembly_result == "PASSED" ~ "PASSED",
    raw_reads_result == "WARNING" & assembly_result == "PASSED" ~ "PASSED",
    raw_reads_result == "FAILED" | assembly_result == "FAILED" ~ "FAILED",
    .default = ""
  )) %>%
  relocate(wgs_id, .before = species_abundance) %>%
  relocate(final_result, .after = wgs_id) %>%
  select(-iso_num)
  
sheet_write(data = overall_result, ss = sheet_id, sheet = "result")

# append googlesheet "result" tab
#sheet_append(ss = sheet_id, sheet = "result", data = overall_result)


write.table(as.matrix(overall_result), paste0("data_files/ARSRL_WGS_QC_report_",batch_code,".csv"), sep = ",", row.names = FALSE)

```




```{r mlst_df ,echo=FALSE, message=FALSE, warning=FALSE}
#rename columns
mlst_df_clean <- df_mlst %>%
  rename("sample_id" = 1,
         "species" = 2,
         "MLST" = 3
         ) %>%
  mutate(
    sample_id = sub(".fna$", "\\1", sample_id),
    sample_id = gsub("-", "_", sample_id)
    ) %>%
  filter(str_detect(sample_id, batch_name_clean))




# Split the string in each row by a semicolon
split_list <- strsplit(mlst_df_clean$ALLELES, split = ";")

# Determine the maximum number of values across all rows
max_cols <- max(sapply(split_list, length))

# Pad shorter lists with NA to make them all the same length
split_list_padded <- lapply(split_list, function(x) {
  length(x) <- max_cols
  return(x)
})

# Convert the list to a matrix and then to a data frame, assigning names
new_cols_df <- as.data.frame(do.call(rbind, split_list_padded))
colnames(new_cols_df) <- paste0("allele", 1:max_cols)

# Combine with the original data frame
mlst_df_clean <- cbind(mlst_df_clean, new_cols_df)



#drop row without species
mlst_df_clean <- subset(mlst_df_clean, mlst_df_clean$species != '-')

#merge wgs_id from wgs_df to mlst_df to set as species value
wgs_species <- subset(wgs_df, select = c("sample_id", "wgs_id"))
mlst_df_clean <- merge(mlst_df_clean,wgs_species,by = "sample_id")

#remove batch name in sample_id
mlst_df_clean$sample_id <- sub("(.*)_[^_]*$", "\\1", mlst_df_clean$sample_id)

#reorder column
mlst_df_clean <- mlst_df_clean %>% 
  relocate(wgs_id, .after = species) %>%
  select(-ALLELES)

#identify salmonella enterica group
mlst_df_clean$species <- ifelse(mlst_df_clean$wgs_id == 'Salmonella enterica', 'salmonella_enterica', ifelse(mlst_df_clean$wgs_id == 'Salmonella typhoidal','salmonella_typhoidal', mlst_df_clean$species))


#not identified sample
#mlst_df_clean[is.na(mlst_df_clean)] <- 'Not Identified'


#list unique species in the dataframe
species_list <- unique(mlst_df_clean$species)

#list unique wgs_id in the dataframe
mlst_organism_list <- unique(mlst_df_clean$wgs_id)

# -- add letter --
mlst_organism_list <- paste0(
  letters[seq_along(mlst_organism_list)],
  ") ",
  mlst_organism_list
)


# -- Add "and" before the last item --
n <- length(mlst_organism_list)

mlst_organism_list[n] <- paste("and", mlst_organism_list[n])

mlst_df_list = c()


#create multiple dataframe for each species
for(i in species_list) {
  #retain first four columns
  mlst_info_df <- subset(mlst_df_clean, select = 1:6)
  
  #remove demogs column
  mlst_result_df <- subset(mlst_df_clean, , select = -c(1:6))


  #use first row data as column names in r
  colnames(mlst_result_df)=mlst_result_df[c(1),]

  #remove special characters and number from column header
  colnames(mlst_result_df) <- gsub("\\s*\\(.*", "", colnames(mlst_result_df))


  # Apply gsub to get text inside the parenthesis
  mlst_result_df <- apply(mlst_result_df, 2, function(x) gsub("\\(([^()]*)\\)|.", "\\1", x, perl=T))
  
  #remove rows with na value
  mlst_result_df <- mlst_result_df[ , colSums(is.na(mlst_result_df))==0]

  # Convert back to data frame
  mlst_result_df <- as.data.frame(mlst_result_df)

  #combine the two dataframe
  mlst_result_df <- cbind(mlst_info_df,mlst_result_df)
  
  #assign new name to dataframe
  nam <- paste("df", i, sep = "_")
  assign(nam, mlst_result_df[mlst_result_df$species==i,])
  
  mlst_df_list <- c(mlst_df_list, nam)
}


```



\fontsize{7}{8}
\selectfont
\captionsetup[table]{labelformat=empty}
\renewcommand{\arraystretch}{1}
\section*{MLST RESULTS}

\small\textbf{\emph{Table 4.} Summary of Multilocus Sequence Typing (MLST) results of \emph{`r paste(mlst_organism_list, collapse = ", ")`}}


```{r mlst_table ,echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
#list existing sample_id in the mlst table
not_included <- wgs_df$wgs_id[!(wgs_df$wgs_id %in% mlst_df_clean$wgs_id)]

  
for(i in mlst_df_list) {
  df <- get(i)
  
  df <- subset(df, select = -c(species) )
  names(df)[names(df) == 'wgs_id'] <- 'species'
  names(df)[names(df) == 'sample_id'] <- 'Sample ID'
  
  col_num <- ncol(df)
  
  #kable(df,  "latex", booktabs = T, align = "c") %>%
    #kable_styling(position = "left") %>%
    #row_spec(0, background = "#D4D4D4",bold = TRUE)

  mlst_table <- kable(df, format='latex', booktabs = TRUE, align = "c", row.names = FALSE) %>%
    kable_styling(latex_options = c("hold_position","scale_down"), position = "left") %>%
    row_spec(0, background = "#D4D4D4",bold = TRUE)%>%
    column_spec(1, width = "3cm") %>%
    column_spec(2, width = "4cm", italic = TRUE) %>%
    column_spec(3, width = "1cm") %>%
    column_spec(4, width = "2cm") %>%
    column_spec(5:col_num, width = "1cm") %>%
    footnote(general ="(-) Not identified",
           general_title = "Legend: ", 
           footnote_as_chunk = T, title_format = c("italic"))


      
  print(mlst_table)

}

```



```{r no_mlst_table ,echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
#list existing sample_id in the mlst table
no_mlst <- sample_name_clean[!(sample_name_clean %in% mlst_df_clean$sample_id)]

if (length(no_mlst) !=0){
  sample_id <- paste(no_mlst, batch_name_clean, sep = "_")
  #sample_id <<- wgs_df$sample_id[(wgs_df$sample_id %in% no_mlst)]
  wgs_id <- wgs_df$wgs_id[(wgs_df$sample_id %in% sample_id)]
  sample_id_clean <- sub("(.*)_[^_]*$", "\\1", sample_id)
  species <- "No MLST result"
  
  df <- data.frame(sample_id_clean, wgs_id, species)
  names(df)[names(df) == 'sample_id_clean'] <- 'sample_id'
  
  no_mlst_table <- kable(df,align = "c", row.names = FALSE) %>%
    kable_styling(position = "left") %>%
    row_spec(0, background = "#D4D4D4",bold = TRUE) %>%
    add_header_above(c("Summary of Sample with No MLST Result" = 3), align = "l", bold = TRUE)
   


  print(no_mlst_table)
  
}



```



\section*{MLST RESULTS SUMMARY}


```{r mlst_summary ,echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
mlst_summary <- data.frame(Species=character(), 
                 MLST=character(), 
                 stringsAsFactors=FALSE) 

for(i in mlst_df_list) {
    df <- get(i)
    
  
    #get unique species value
    species <- unique(df$wgs_id)
    
    #remove from list all value with "quasi"
    species <- species[!species %in% grep("quasi", species, value = T)]

    #Summarize and count MLST results
    mlst_count_df <- df %>% group_by(wgs_id,MLST) %>% summarise(count = sum(!is.na(MLST)))
    mlst_count_df$mlst_count <- paste0(mlst_count_df$MLST, " (n= ", mlst_count_df$count, " )")

    #Remove empty MLST result
    #mlst_count_df  <- subset(mlst_count_df , mlst_count_df $MLST !='-')
    mlst_count_summary  <- subset(mlst_count_df , select = c(wgs_id, mlst_count))
    mlst_summary_df <- mlst_count_summary %>% 
      group_by(wgs_id) %>% 
      mutate(mlst_count = paste0(mlst_count, collapse = ", ")) %>%
      distinct()
    
    #assign new name to dataframe
    #nam <- paste("mlst", i, sep = "_")
    #assign(nam, mlst_summary)
    
    mlst_summary <- rbind(mlst_summary,mlst_summary_df)
    
    #add new data to exisiting dataframe
    #mlst_summary[nrow(mlst_summary) + 1,] = c(species, paste0(mlst_count_df$mlst_count, collapse=", "))
  
}

#Print table
 mlst_summary_table <- kable(mlst_summary, format='latex', booktabs = TRUE, align = "l", row.names = FALSE) %>%
    kable_styling(position = "left") %>%
    column_spec(1, width = "6cm", italic = TRUE) %>%
    column_spec(2, width = "10cm") %>%
    row_spec(0, background = "#D4D4D4",bold = TRUE) %>%
    footnote(general ="(-) Not identified",
           general_title = "Legend: ", 
           footnote_as_chunk = T, title_format = c("italic"))
      
  print(mlst_summary_table)
 
```

\newpage
\fontsize{7}{8}
\selectfont
\captionsetup[table]{labelformat=empty}
\renewcommand{\arraystretch}{1.2}




```{r amr_genes ,echo=FALSE, message=FALSE, warning=FALSE}
df_amrfinderplus1 <- df_amrfinderplus

#checkm2 column names to lower
names(df_amrfinderplus1) <- tolower(names(df_amrfinderplus1))


#Change Name to sample_id
names(df_amrfinderplus1)[names(df_amrfinderplus1) == 'name'] <- 'sample_id'
names(df_amrfinderplus1)[names(df_amrfinderplus1) == 'element symbol'] <- 'element_symbol'

#change "-" to "_" in sample_id column
df_amrfinderplus1$sample_id <- gsub("-", "_", df_amrfinderplus1$sample_id, fixed=TRUE)


#merge wgs_id from wgs_df to mlst_df to set as species value
df_amrfinderplus1 <- merge(df_amrfinderplus1,wgs_species,by = "sample_id")

#remove "quasi" to wgs_id column
df_amrfinderplus1$wgs_id <- gsub("quasi", "", df_amrfinderplus1$wgs_id, fixed=TRUE)



# Filter rows where 'name' contains the word "sample"
df_amrfinderplus1 <- df_amrfinderplus1[grepl(batch_name_clean, df_amrfinderplus1$sample_id), ]

#remove batch name in sample_id
df_amrfinderplus1$sample_id <- sub("(.*)_[^_]*$", "\\1", df_amrfinderplus1$sample_id)

amr_genes_species <- unique(df_amrfinderplus1$wgs_id)

# -- add letter --
amr_genes_list <- paste0(
  letters[seq_along(amr_genes_species)],
  ") ",
  amr_genes_species
)


# -- Add "and" before the last item --
n <- length(amr_genes_list)

amr_genes_list[n] <- paste("and", amr_genes_list[n])

#vector to hold list of amr gene datables
amr_gene_df_list = c()


for(i in amr_genes_species) {
  df <- subset(df_amrfinderplus1, df_amrfinderplus1$wgs_id == i)
  sample_id <- c(df$sample_id)
  
  #create new dataframe to duplicate df_amrfinderplus1
  amr_genes_df <- df_amrfinderplus1
  
  #change "NA" to " " in subclass column
  amr_genes_df$subclass <- gsub("NA", "", amr_genes_df$subclass, fixed=TRUE)
  
  amr_genes_df$Newcols <- paste(amr_genes_df$type,amr_genes_df$subclass, sep = " ")
  
  #change "/" to "/ " in sample_id column
  amr_genes_df$Newcols <- gsub("/", "/ ", amr_genes_df$Newcols, fixed=TRUE)
  
  #subset dataframe to retain needed columns only
  amr_genes_df <- subset(amr_genes_df, select = c(sample_id,Newcols,element_symbol))
  
  
  #convert multiple rows with same Newcols value to single row
  amr_genes_df <- amr_genes_df %>%
    group_by(Newcols, sample_id) %>%
    summarize(Concat = paste0(element_symbol, collapse = ", "))
  
  #Create new dataframe to summarize the result
  amr_genes_df <- amr_genes_df  %>%
    pivot_wider(names_from = Newcols,
                values_from = Concat)
 #assign new name to dataframe
  nam <- paste("amr", i, sep = "_")
  assign(nam, amr_genes_df[amr_genes_df$sample_id %in% sample_id, ])
  
  amr_gene_df_list <- c(amr_gene_df_list, nam)

}


```


\fontsize{7}{8}
\selectfont
\captionsetup[table]{labelformat=empty}
\renewcommand{\arraystretch}{1.2}
\section*{AMR PREDICTION RESULTS}


\small\textbf{\emph{Table 5.} Summary of AMR genes present among of \emph{`r paste(amr_genes_list, collapse = ", ")`}}


```{r amr_genes_table ,echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
for (idx in seq_along(amr_gene_df_list)) {
  i <- amr_gene_df_list[idx]
  df <- get(i)
  
  wgs_species_1 <- wgs_species
  
  #remove batch name in sample_id
  wgs_species_1$sample_id <- sub("(.*)_[^_]*$", "\\1", wgs_species_1$sample_id)

  # Merge wgs_id from wgs_df to amr_gene to set as species value
  df <- merge(df, wgs_species_1, by = "sample_id")
  
  # Change wgs_id to species
  names(df)[names(df) == 'wgs_id'] <- 'species'
  names(df)[names(df) == 'sample_id'] <- 'Sample ID'
  
  # Reorder column
  df <- df %>% relocate(species, .after = `Sample ID`)
  
  # Remove columns with all NA values
  df <- df[, colSums(is.na(df)) < nrow(df)]
  
  # Get unique species value
  species <- unique(df$species)
  species_caption <- paste("\\\\textit{", species, "}", sep = "")
  species_caption <- species_caption[1]
  
  # -- add letter --
  species_caption <- paste0(
    letters[idx],
    ") ",
    species_caption
  )
  
  # Subset dataframe to retain needed columns only
  df <- subset(df, select = -c(species))
  
  # Parameters for splitting
  max_rows_per_table <- 6
  max_cols_per_table <- 6  # Maximum columns per table (excluding sample_id)
  
  # Get the column names (excluding sample_id)
  all_cols <- setdiff(colnames(df), "Sample ID")
  
  # Calculate number of column groups needed
  num_col_groups <- ceiling(length(all_cols) / max_cols_per_table)
  
  # Split rows first
  num_row_groups <- ceiling(nrow(df) / max_rows_per_table)
  
  for(row_group in 1:num_row_groups) {
    # Determine row range for this group
    start_row <- (row_group-1) * max_rows_per_table + 1
    end_row <- min(row_group * max_rows_per_table, nrow(df))
    df_row_subset <- df[start_row:end_row, , drop = FALSE]
    
    # Now split columns for this row group
    for(col_group in 1:num_col_groups) {
      # Determine column indices for this group
      start_col <- (col_group-1) * max_cols_per_table + 1
      end_col <- min(col_group * max_cols_per_table, length(all_cols))
      cols_to_keep <- c("Sample ID", all_cols[start_col:end_col])
      
      df_subset <- df_row_subset[, cols_to_keep, drop = FALSE]
      
      # Create table
      amr_gene_table <- knitr::kable(df_subset, format='latex', booktabs = TRUE, align = "c") %>%
        kable_styling(
          position = "left",
          latex_options = c("HOLD_position", "scale_down"), 
          font_size = 7
        ) %>%
        row_spec(0, background = "#D4D4D4", bold = TRUE) %>%
        column_spec(2:ncol(df_subset), width = "3cm") %>%
        add_header_above(setNames(ncol(df_subset), 
                                paste0(species_caption, 
                                      " (Part ", row_group, ".", col_group, ")")), 
                        bold = TRUE, escape = FALSE, align = "l")
      
      print(amr_gene_table)
      
      # Add vertical space between column groups (but not after last one)
      if(col_group < num_col_groups) cat("\n\n\\vspace{5mm}\n\n")
    }
    
    # Add more vertical space between row groups (but not after last one)
    if(row_group < num_row_groups) cat("\n\n\\vspace{10mm}\n\n")
  }
}
```






```{r no_amr_genes_table ,echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
#list existing sample_id in the mlst table
no_amr <- wgs_df$wgs_id[!(df_amrfinderplus1$wgs_id %in% wgs_df$wgs_id)]

#remove if with quasi value
no_amr <-no_amr[!grepl('quasi', no_amr)]

if (length(no_amr) !=0){
  sample_id <- wgs_df$sample_id[!(wgs_df$sample_id %in% paste(df_amrfinderplus1$sample_name, batch_name_clean, sep = "_"))]
  wgs_id <- no_amr
  remarks <- "No AMR determinant detected"
  
  df <- data.frame(sample_id, wgs_id, remarks)
  
  no_amr_table <- kable(df,align = "c", row.names = FALSE) %>%
    kable_styling(position = "left") %>%
    row_spec(0, background = "#D4D4D4",bold = TRUE)
   


  print(no_amr_table)
  
}

```

\end{landscape}
